Introducing ICBe: Very High Recall and Precision Event Extraction from Narratives about International Crises

### 2022-07-25 replication_paper/arxiv_draft/paper.tex
arxiv draft

\documentclass[9pt,twocolumn,twoside,]{pnas-new}

% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

@@

% From pandoc table feature
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

@@

\usepackage{xr} \externaldocument{appendix}
\usepackage[utf8]{inputenc}
\usepackage{pifont}
\usepackage{newunicodechar}
\newunicodechar{✓}{\ding{51}}
\newunicodechar{✗}{\ding{55}}
\usepackage{array}
\usepackage{ctable}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{titlesec}
\usepackage[parfill]{parskip}
\usepackage{makecell}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{cellspace}
\setlength\cellspacetoplimit{0.8ex}
\renewcommand{\arraystretch}{0.8}
\AtBeginEnvironment{tabular}{\singlespacing}
\AtBeginEnvironment{lltable}{\singlespacing}
\usepackage{subcaption}
\usepackage{atbegshi}% http://ctan.org/pkg/atbegshi
\usepackage{float}
\usepackage[algo2e]{algorithm2e}
\usepackage{fontspec}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}

\templatetype{pnasresearcharticle}  % Choose template

\author[a,1,2]{Rex W. Douglass}
\author[a]{Thomas Leo Scherer}
\author[b]{J. Andrés Gannon}
\author[a]{Erik Gartzke}
\author[c]{Jon Lindsay}
\author[d]{Shannon Carcelli}
\author[d]{Jonathan Wilkenfeld}
\author[e]{David M. Quinn}
\author[f]{Catherine Aiken}
\author[g]{Jose Miguel Cabezas Navarro}
\author[d]{Neil Lund}
\author[h]{Egle Murauskaite}
\author[h]{Diana Partridge}

  \affil[a]{Department of Political Science, University of California, San Diego, CA, USA.}
  \affil[b]{Belfer Center for Science and International Affairs, Harvard Kennedy School, MA, USA.}
  \affil[c]{School of Cybersecurity and Privacy \textbar{} Sam Nunn School of International Affairs, Georgia Institute of Technology, GA, USA.}
  \affil[d]{Department of Government and Politics, University of Maryland, College Park, MD, USA.}
  \affil[e]{Faculty Specialist, National Consortium for the Study of Terrorism and Responses to Terrorism (START), University of Maryland, College Park, MD, USA.}
  \affil[f]{Edmund A. Walsh School of Foreign Service, Georgetown University, Washington, DC, USA.}
  \affil[g]{Society and Health Research Center, Universidad Mayor, Santiago, Chile.}
  \affil[h]{ICONS Project // START, University of Maryland, College Park, MD, USA.}
  \affil[i]{ICONS Project // START, University of Maryland, College Park, MD, USA.}

% Please give the surname of the lead author for the running footer
\leadauthor{Douglass}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{Countries routinely face crises that risk escalating into full scale war but we do not have systematic measurements of the progression of past crises and what moves and counter moves led to or helped avoid war. Instead policy makers typically rely on one or two historical analogies, chosen through ad hoc selection criteria, and described in unsystematic terms. This paper introduces a new scientific approach to measuring the step by step moves of international crises over the last hundred years, combining subject expertise with state of the art natural language processing and machine learning methods. It serves as a guide for constructing and evaluating large scale measurement collection in the social sciences.}

\authorcontributions{Conceptualization: R.W.D., E.G., J.L.; Methodology: R.W.D., T.L.S.; Software: R.W.D.; Validation: R.W.D., T.L.S.; Formal Analysis: R.W.D., T.L.S.; Investigation: S.C., R.W.D., J.A.G., C.K., N.L., E.M., J.M.C.N., D.P., D.Q., J.W.; Data Curation: R.W.D., D.Q., T.L.S., J.W.; Writing - Original Draft: R.W.D., T.L.S.; Writing - Review \& Editing: R.W.D., J.A.G., E.G., T.L.S.;
Visualization: R.W.D., T.L.S.; Supervision: E.G.; Project Administration: S.C., R.W.D., J.A.G., D.Q., T.L.S., J.W.; Funding Acquisition: E.G., J.L.}

\authordeclaration{Please declare any conflict of interest here.}

\correspondingauthor{\textsuperscript{2} To whom correspondence should be addressed. E-mail: \href{mailto:rexdouglass@gmail.com}{\nolinkurl{rexdouglass@gmail.com}}}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
 \keywords{  Diplomacy |  War |  Crises |  International Affairs |  Computational Social Science  }

\begin{abstract}
How do international crises unfold? We conceptualize of international relations as a strategic chess game between adversaries and develop a systematic way to measure pieces, moves, and gambits accurately and consistently over a hundred years of history. We introduce a new ontology and dataset of international events called ICBe based on very high-quality corpus of narratives from the International Crisis Behavior (ICB) Project. We demonstrate that ICBe has higher coverage, recall, and precision than existing state of the art datasets and conduct two detailed case studies of the Cuban Missile Crisis (1962) and Crimea-Donbas Crisis (2014). We further introduce two new event visualizations (event icongraphy and crisis maps), an automated benchmark for measuring event recall using natural language processing (sythnetic narratives), and an ontology reconstruction task for objectively measuring event precision. We make the data, replication material, and visualizations of every historical episode available at a companion website www.crisisevents.org.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.

\acknow{We thank the ICB Project and its directors and contributors for their foundational work and their help with this effort. We make special acknowledgment of Michael Brecher for helping found the ICB project in 1975, creating a resource that continues to spark new insights to this day. We thank the many undergraduate coders for their patience and dedication. Thanks to the Center for Peace and Security Studies and its membership for comments. Special thanks to Rebecca Cordell, Philip Schrodt, Zachary Steinert-Threlkeld, and Zhanna Terechshenko for generous feedback. Thank you to the cPASS research assistants that contributed to this project: Helen Chung, Daman Heer, Syeda ShahBano Ijaz, Anthony Limon, Erin Ling, Ari Michelson, Prithviraj Pahwa, Gianna Pedro, Tobias Stodiek, Yiyi `Effie' Sun, Erin Werner, Lisa Yen, and Ruixuan Zhang. This project was supported by a grant from the Office of Naval Research {[}N00014-19-1-2491{]} and benefited from the Charles Koch Foundation's support for the Center for Peace and Security Studies.}

If we could record every important interaction between countries in all of diplomacy, military conflict, and international political economy, how much unique information would this chronicle amount to, and how surprised would we be to see something new? In other words, what is the entropy of international relations? This record could in principle be unbounded, but the central conceit of social science is that there are structural regularities that limit what actors can do, their best options, and even which actors are likely to survive (1, 2). If so, then these events can be systematically measured, and accordingly, massive effort is expended in social science attempting to record these regularities.\footnote{See work on crises (3, 4), militarized disputes (5--7), wars (8, 9), organized violence (10, 11), political violence (12), sanctions (13), trade (14), and international agreements (15--17), dispute resolution (17, 18), and diplomacy (19, 20).} With improvements in natural language processing, more open-ended efforts have sought to capture entire unstructured streams of international events drawn from news reports.\footnote{See (21); (22); (23); (24); (25); (26). On event-extraction from images and social-media see (27) and (28).} How close these efforts are to accurately measuring all or even most of what is essential in international relations is an open empirical question, one for which we provide new evidence here.

@@

The first axis (rows) represents the types of information we expect to find in IR and forms the basis for our proposed ontology. We employ a metaphor of a chess game, with players (polities, rebel groups, IGOs, etc.), pieces (military platforms, civilians, domains), and behaviors (think, say, do). Precise sequencing is required to capture gambits (sequences of moves) and outcomes (victory, defeat, peace, etc.), while precise geo-coding is required to understand the chessboard (medium of conflict). We find 464 actors and 117 different behaviors, and provide a full codebook in the online material.\footnote{See the Github Repository \href{https://urldefense.com/v3/__https://github.com/CenterForPeaceAndSecurityStudies/ICBEventData__;!!Mih3wA!WxDJtEczKfxGTh0S2Krunap8ReymFEL5iTWaSfOHeqlSdyfRx77zmjBSWO1OAm13$}{ICBEventData}.}

@@

\multicolumn{2}{!{\color[HTML]{000000}\vrule width 2pt}>{\centering}p{\dimexpr 0.2in+2\tabcolsep+1\arrayrulewidth}}{} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{EFEFEF}\centering}p{\dimexpr 3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{7}{4}\selectfont{\textcolor[HTML]{000000}{N}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{EFEFEF}\centering}p{\dimexpr 0.25in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{7}{4}\selectfont{\textcolor[HTML]{000000}{15k}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\centering}p{\dimexpr 0.25in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{7}{4}\selectfont{\textcolor[HTML]{000000}{8.5M}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\centering}p{\dimexpr 0.25in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{7}{4}\selectfont{\textcolor[HTML]{000000}{28.4M}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\centering}p{\dimexpr 0.25in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{7}{4}\selectfont{\textcolor[HTML]{000000}{17.5M}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\centering}p{\dimexpr 0.25in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{7}{4}\selectfont{\textcolor[HTML]{000000}{9.6K}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\centering}p{\dimexpr 0.25in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{7}{4}\selectfont{\textcolor[HTML]{000000}{128k}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{EFEFEF}\centering}p{\dimexpr 0.25in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{7}{4}\selectfont{\textcolor[HTML]{000000}{1K}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\centering}p{\dimexpr 0.25in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{7}{4}\selectfont{\textcolor[HTML]{000000}{5.9K}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\centering}p{\dimexpr 0.25in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 2pt}}{\fontsize{7}{4}\selectfont{\textcolor[HTML]{000000}{1K}}} \\

@@

Each crisis was typically assigned to 2 expert coders and 2 novice coders with an additional tie-breaking expert coder assigned to sentences with high disagreement.\footnote{Expert coders were graduate students or postgraduates who collaboratively developed the ontology and documentation for the codebook. Undergraduate coders were students who engaged in classroom workshops.} For the purposes of measuring intercoder agreement and consensus, we temporarily disaggregate the unit of analysis to the Coder-Crisis-Sentence-Tag (n=972,545), where a tag is any unique piece of information a coder can associate with a sentence such as an actor, date, behavior, etc. We then aggregate those tags into final events (n=18,309), using a consensus procedure (SI Appendix, Algorithm B2) that requires a tag to have been chosen by at least one expert coder and either a majority of expert or novice coders. This screens noisy tags that no expert considered possible but leverages novice knowledge to tie-break between equally plausible tags chosen by experts.

@@

We evaluate the internal validity of the coding process in several ways. For every tag applied we calculate the observed intercoder agreement as the percent of other coders who also applied that same tag (SI Appendix, Fig. B3). Across all concepts, the Top 1 Tag Agreement was low among novices (31\%), moderate for experts (65\%), and high (74\%) following the consensus screening procedure.

@@

\clearpage
\onecolumn
\begin{figure}[H]
\caption{ICBe event codings in comparison to Semantic Embeddings from source sentences\label{fig:semantic_embeddings}}
\centering{\includegraphics[width=17cm]{../figures/p_semantic_embeddings_dendro.png}}
\textit{Notes: Dots represent individual ICB narrative sentences, as embeded by the Paraphrase-MPNET-base-v2 large language model and flattened into two dimensions with UMAP. Text labels reflect individual leaves of the ICBe ontology, and colors represent intermediate branches of the ontology. Label placement is the median of all of the sentences that tag was applied to by the coders. The dendrogram shows hiearchical clustering of the tags. If ICBe precision is high, the sentences tags were applied to ought to say similar things, and the intended shape of the ontology ought to visually recognizable. }
\end{figure}
\clearpage
\twocolumn

Finally, how does ICBe's precision compare to the existing state of the art? The crisis-maps reveal the episode level datasets like MIDs or the original ICB are too sparse and vague to reconstruct the structure of the crisis. On the other end of the spectrum, the high recall dictionary based event datasets like Terrier and ICEWs produce so many noisy events (several hundreds thousands) that even with heavy filtering their crisis maps are completely unintelligible. Further, because of copyright issues, none of these datasets directly provide the original text spans making event level precision difficult to verify.

@@

\clearpage
\onecolumn
\begin{figure}[H]
\caption{ ICEWs Events by Day by Type during the Crimea-Donbas Crisis \label{fig:p_precision_icews}}
\centering{\includegraphics[width=17cm]{../figures/p_precision_icews.png}}
\textit{Notes:Unit of analysis is the Dyad-Day. Edges <-> indicates undirected dyad and -> indicates directed dyad. Top 10 most active dyads per category shown. Red text shows events from the synthetic narrative relative to that event category. Blue bars indicate an event recorded by ICEWs for that dyad on that day. }
\end{figure}
\clearpage
\twocolumn

@@

We investigated event extraction from narratives describing key historical episodes in international relations. We synthesized a prior belief about the latent unobserved phenomena that drive these events in international relations and proposed a mapping to observable concepts that enter into the observed historical record. We designed an ontology with high coverage over those concepts and developed a training procedure and technical stack for human coding of historical texts. Multiple validity checks find the resulting codings have high internal validity (e.g.~intercoder agreement) and external validity (i.e.~matching source material in both micro-details at the sentence level and macro-details spanning full historical episodes). Further, these codings perform much better in terms of recall, precision, coverage, and overall coherence in capturing these historical episodes than existing event systems used in international relations.

@@

\showmatmethods
\showacknow
\pnasbreak

@@

% Bibliography
% \bibliography{pnas-sample}

>>>>

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}

\author{
  }

@@

@@

\usepackage[utf8]{inputenc}
\usepackage{pifont}
\usepackage{newunicodechar}
\newunicodechar{✓}{\ding{51}}
\newunicodechar{✗}{\ding{55}}
\usepackage{array}
\usepackage{ctable}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{titlesec}
\usepackage[parfill]{parskip}
\usepackage{makecell}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{cellspace}
\setlength\cellspacetoplimit{0.8ex}
\renewcommand{\arraystretch}{0.8}
\AtBeginEnvironment{tabular}{\singlespacing}
\AtBeginEnvironment{lltable}{\singlespacing}
\usepackage{subcaption}
\usepackage{atbegshi}% http://ctan.org/pkg/atbegshi
\usepackage{float}
\usepackage[algo2e]{algorithm2e}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}
\begin{document}
\maketitle

\begin{abstract}
How do international crises unfold? We conceptualize of international
relations as a strategic chess game between adversaries and develop a
systematic way to measure pieces, moves, and gambits accurately and
consistently over a hundred years of history. We introduce a new
ontology and dataset of international events called ICBe based on a very
high-quality corpus of narratives from the International Crisis Behavior
(ICB) Project. We demonstrate that ICBe has higher coverage, recall, and
precision than existing state of the art datasets and conduct two
detailed case studies of the Cuban Missile Crisis (1962) and
Crimea-Donbas Crisis (2014). We further introduce two new event
visualizations (event icongraphy and crisis maps), an automated
benchmark for measuring event recall using natural language processing
(sythnetic narratives), and an ontology reconstruction task for
objectively measuring event precision. We make the data, replication
material, and visualizations of every historical episode available at a
companion website www.crisisevents.org.
\end{abstract}

\keywords{
    Diplomacy
   \and
    War
   \and
    Crises
   \and
    International Affairs
   \and
    Computational Social Science
  }

\twocolumn

If we could record every important interaction between countries in all
of diplomacy, military conflict, and international political economy,
how much unique information would this chronicle amount to, and how
surprised would we be to see something new? In other words, what is the
entropy of international relations? This record could in principle be
unbounded, but the central conceit of social science is that there are
structural regularities that limit what actors can do, their best
options, and even which actors are likely to survive (1, 2). If so, then
these events can be systematically measured, and accordingly, massive
effort is expended in social science attempting to record these
regularities.\footnote{See work on crises (3, 4), militarized disputes
  (5--7), wars (8, 9), organized violence (10, 11), political violence
  (12), sanctions (13), trade (14), and international agreements
  (15--17), dispute resolution (17, 18), and diplomacy (19, 20).} Thanks
to improvements in natural language processing, more open-ended efforts
have begun to capture entire unstructured streams of events from
international news reports.\footnote{See (21); (22); (23); (24); (25);
  (26). On event-extraction from images and social-media see (27) and
  (28).} How close these efforts are to accurately measuring all or even
most of what is essential in international relations is an open
empirical question, one for which we provide new evidence here.

@@

The first axis (rows) represents the types of information we expect to
find in IR and forms the basis for our proposed ontology. We employ a
metaphor of a chess game, with players (polities, rebel groups, IGOs,
etc.), pieces (military platforms, civilians, domains), and behaviors
(think, say, do). Precise sequencing is required to capture gambits
(sequences of moves) and outcomes (victory, defeat, peace, etc.), while
precise geo-coding is required to understand the chessboard (medium of
conflict). We find 472 actors and 117 different behaviors, and provide a
full codebook in the online material.\footnote{See the Github Repository
  \href{https://urldefense.com/v3/__https://github.com/CenterForPeaceAndSecurityStudies/ICBEventData__;!!Mih3wA!WxDJtEczKfxGTh0S2Krunap8ReymFEL5iTWaSfOHeqlSdyfRx77zmjBSWO1OAm13$}{ICBEventData}.}

@@

\multicolumn{2}{!{\color[HTML]{000000}\vrule width 2pt}>{\centering}p{\dimexpr 0.2in+2\tabcolsep+1\arrayrulewidth}}{} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{EFEFEF}\centering}p{\dimexpr 3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{7}{4}\selectfont{\textcolor[HTML]{000000}{N}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{EFEFEF}\centering}p{\dimexpr 0.25in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{7}{4}\selectfont{\textcolor[HTML]{000000}{32k}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\centering}p{\dimexpr 0.25in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{7}{4}\selectfont{\textcolor[HTML]{000000}{8.5M}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\centering}p{\dimexpr 0.25in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{7}{4}\selectfont{\textcolor[HTML]{000000}{28.4M}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\centering}p{\dimexpr 0.25in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{7}{4}\selectfont{\textcolor[HTML]{000000}{17.5M}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\centering}p{\dimexpr 0.25in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{7}{4}\selectfont{\textcolor[HTML]{000000}{9.6K}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\centering}p{\dimexpr 0.25in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{7}{4}\selectfont{\textcolor[HTML]{000000}{128k}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{EFEFEF}\centering}p{\dimexpr 0.25in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{7}{4}\selectfont{\textcolor[HTML]{000000}{1K}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\centering}p{\dimexpr 0.25in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{7}{4}\selectfont{\textcolor[HTML]{000000}{5.9K}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\centering}p{\dimexpr 0.25in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 2pt}}{\fontsize{7}{4}\selectfont{\textcolor[HTML]{000000}{1K}}} \\

@@

Each crisis was typically assigned to 2 expert coders and 2 novice
coders with an additional tie-breaking expert coder assigned to
sentences with high disagreement.\footnote{Expert coders were graduate
  students or postgraduates who collaboratively developed the ontology
  and documentation for the codebook. Undergraduate coders were students
  who engaged in classroom workshops.} For the purposes of measuring
intercoder agreement and consensus, we temporarily disaggregate the unit
of analysis to the Coder-Crisis-Sentence-Tag (n=993,740), where a tag is
any unique piece of information a coder can associate with a sentence
such as an actor, date, behavior, etc. We then aggregate those tags into
final events (n=18,783), using a consensus procedure (SI Appendix,
Algorithm B2) that requires a tag to have been chosen by at least one
expert coder and either a majority of expert or novice coders. This
screens noisy tags that no expert considered possible but leverages
novice knowledge to tie-break between equally plausible tags chosen by
experts.

@@

We evaluate the internal validity of the coding process in several ways.
For every tag applied we calculate the observed intercoder agreement as
the percent of other coders who also applied that same tag (SI Appendix,
Fig. B3). Across all concepts, the Top 1 Tag Agreement was low among
novices (31\%), moderate for experts (65\%), and high (73\%) following
the consensus screening procedure.

@@

\clearpage
\onecolumn
\begin{figure}[H]
\caption{ICBe event codings in comparison to Semantic Embeddings from source sentences\label{fig:semantic_embeddings}}
\centering{\includegraphics[width=17cm]{./p_semantic_embeddings_dendro.png}}
\textit{Notes: Dots represent individual ICB narrative sentences, as embeded by the Paraphrase-MPNET-base-v2 large language model and flattened into two dimensions with UMAP. Text labels reflect individual leaves of the ICBe ontology, and colors represent intermediate branches of the ontology. Label placement is the median of all of the sentences that tag was applied to by the coders. The dendrogram shows hiearchical clustering of the tags. If ICBe precision is high, the sentences tags were applied to ought to say similar things, and the intended shape of the ontology ought to be visually recognizable. }
\end{figure}
\clearpage
\twocolumn

Finally, how does ICBe's precision compare to the existing state of the
art? The crisis-maps reveal the episode level datasets like MIDs or the
original ICB are too sparse and vague to reconstruct the structure of
the crisis (SI Appendix Figure D3 and D4). On the other end of the
spectrum, the high recall dictionary based event datasets like Terrier
and ICEWs produce so many noisy events (several hundreds thousands) that
even with heavy filtering their crisis maps are completely
unintelligible. Further, because of copyright issues, none of these
datasets directly provide the original text spans making event level
precision difficult to verify.

@@

\clearpage
\onecolumn
\begin{figure}[H]
\caption{ ICEWs Events by Day by Type during the Crimea-Donbas Crisis \label{fig:p_precision_icews}}
\centering{\includegraphics[width=17cm]{./p_precision_icews.png}}
\textit{Notes:Unit of analysis is the Dyad-Day. Edges <-> indicates undirected dyad and -> indicates directed dyad. Top 10 most active dyads per category shown. Red text shows events from the synthetic narrative relative to that event category. Blue bars indicate an event recorded by ICEWs for that dyad on that day. }
\end{figure}
\clearpage
\twocolumn

@@

We investigated event abstraction from narratives describing key
historical episodes in international relations. We synthesized a prior
belief about the latent unobserved phenomena that drive these events in
international relations and proposed a mapping to observable concepts
that enter into the observed historical record. We designed an ontology
with high coverage over those concepts and developed a training
procedure and technical stack for human coding of historical texts.
Multiple validity checks find the resulting codings have high internal
validity (e.g.~intercoder agreement) and external validity
(i.e.~matching source material in both micro-details at the sentence
level and macro-details spanning full historical episodes). Further,
these codings perform much better in terms of recall, precision,
coverage, and overall coherence in capturing these historical episodes
than existing event systems used in international relations.

@@

@@

\bibliographystyle{unsrt}
\bibliography{paper.bib}
